Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-05 paper_jasa.tex
finished smoothing out conclusions

One of the most critical steps in visual test would be select an appropriate test statistic for the hypothesis of interest.  A good test statistic should show a very specific pattern if the null hypothesis is not true. A small difference in color template or orientation of the plot could make a difference. This is the part of our ongoing and future research.

Even though a single observer can make a decision we recommend having multiple observer to evaluate a lineup to obtain a definitive $p$-value to make a decision confidently.

The experimental subjects that we recruited came from a various level of education and not necessarily well aware of statistical plots. But still we achieve the power of visual test comparable to the conventional test. This suggests that visual statistical technique could be a smart tool for high school students to get to know the first course of statistical inference. The major advantage of visual statical inference is its intuitive concept of testing technique which is easy to be grasped and fun to learn.

@@

When there is no conventional test available, we can conveniently use visual testing procedure due to its flexibility of having no assumption restrictions. Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$). The direction of future research would be to study that and apply the methods in those areas.

The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods. We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.  The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used. For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots. Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.

\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice. Along the lines of the Army tests that Susan brought to the working group. \blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects. Thus anyone from any background could participate the experiment. It would be interesting to see how well the trained subjects performs. Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects. One of our future work would be to examine this in details. For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show. After being trained they may have to pass the qualification test before participating the experiment. Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.

Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}

In practice there is no correct answer. In these experiments we controlled the effects that observers saw, so there was a correct choice in each case. In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it. And subjects would provide open-ended answers for their choice of plot.

What are your suggestions for future experiments, and next steps of the research?

>>>>

One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.  A good test statistic is expected to reveal structure if the null hypothesis is not true. Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference. This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices. %This is the part of our ongoing and future research.

Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers. This ensures the ability to obtain a good estimate of the $p$-value for the test, and more confidence in the decision.

Experimental subjects in the simulation study came from various level of education, both genders and many different ages. Many subjects had little background in statistics, and were not necessarily aware of statistical plots. Yet the results were good. We expect that visual inference could a useful tool for high school students to learn about randomness as a primer to statistical thinking. A major advantage of visual statisticcal inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it is may make statistics fun to learn.

No restrictions were placed on Turk workers, in terms of abilities. It would be interesting to see how well the trained subjects performs. Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects. One of our future work would be to examine this in details. For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show. After being trained they may have to pass the qualification test before participating the experiment. Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.

@@

When there is no conventional test available, visual testing procedures can be used due to its few assumptions and non-parametric nature. We have successfully used the procedures informally for local projects. It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no exising gauges of statistical significance. The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance. If the null hypothesis is rejected, it is generally that we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection. Follow-up questions on the reasons provide qualitative insight. Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.

%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$). The direction of future research would be to study that and apply the methods in those areas.

%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods. We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.  The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used. For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots. Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.

%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice. Along the lines of the Army tests that Susan brought to the working group. \blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects. Thus anyone from any background could participate the experiment. It would be interesting to see how well the trained subjects performs. Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects. One of our future work would be to examine this in details. For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show. After being trained they may have to pass the qualification test before participating the experiment. Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.

%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}

%In practice there is no correct answer. In these experiments we controlled the effects that observers saw, so there was a correct choice in each case. In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it. And subjects would provide open-ended answers for their choice of plot.

%What are your suggestions for future experiments, and next steps of the research?
